---
title: "OverviewAutoProcessedData_PrepManualAnalysis__Sara2025"
output: html_document
date: "2025-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sara's Masters thesis data prep and innitial overview

Target sites that only Sara is looking at so far: 
CM-18, CM-19, CM-26, CM-28, CM-39, CM-40

Target sites that Maris is going to analyze: 
CM-20, CM-22, CM-28, CM-49 

Target autoids - top priority: 
BARBAR, NYCNOC, PIPNAT 

Potential additional autoids: 
NoID, Noise, EPTNIL... 

Need to compile a dataset that Sara can feasibly analyze for her MSc: 
approximately 20,000 files

I need to look at all the data from these sites and see what subset of data is feasible for her to work with.


## Prepare workspace
```{r}
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
library(janitor)
library(renv)
library(stringr)
library(beepr)
library(randomcoloR)
library(wesanderson)
library(leaflegend)
library(osmdata)
library(MetBrewer)
library(colorBlindness)
library(colorblindcheck)
library(MoMAColors)

getwd()
#  "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/GitHubLink/NocturBats"

## Setup output directory 
output <-"C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Analyses/Outputs/Reed" # where you want to save your data

file.name <- "OverviewProcessedBatData_Sara2025"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today
# For example: 
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Analyses/Outputs/Reed/OverviewProcessedBatData_Sara2025_2025-01-09"

```

## Read in the processed data and look for potential errors 
```{r}

# Sites for Sara
# CM-18, CM-19, CM-28, CM-39, CM-40, CM-47
# AutoIDs for Sara: 
## PIPNAT, NYCNOC 

inputCM18 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-18/WAV/KPRO_V1/id.csv") 
# 29749 obs of 44 vars
# Add a site column 
inputCM18$site <- "CM-18"

inputCM19 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-19/WAV/KPRO_V1/id.csv") 
# 16684 obs of 44 vars
# Add a site column 
inputCM19$site <- "CM-19"

# CM-26
inputCM26 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-26/WAV/KPRO_V1/id.csv") 
# 33595 obs of 44 vars
# Add a site column 
inputCM26$site <- "CM-26"

inputCM39 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-39/WAV/KPRO_V1/id.csv") 
# 14716 obs of 44 vars
# Add a site column 
inputCM39$site <- "CM-39"

inputCM40 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-40/WAV/KPRO_V1/id.csv") 
# 25388 obs of 44 vars
# Add a site column 
inputCM40$site <- "CM-40"


## These are sites that Maris plans to manually analyze 

## but we we should still take a look at them and see what files are there

# CM-20
inputCM20 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-20/WAV/KPRO_V1/id.csv") 
# 19711 obs of 44 vars
# Add a site column 
inputCM20$site <- "CM-20"

# CM-22
inputCM22 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-22/WAV/KPRO_V1/id.csv") 
# 44375 obs of 44 vars
# Add a site column 
inputCM22$site <- "CM-22"

# CM-28
inputCM28 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-28/WAV/KPRO_V1/id.csv") 
# 54918 obs of 44 vars
# Add a site column 
inputCM28$site <- "CM-28"

test <- inputCM28 %>% mutate(autoid = factor(`AUTO ID*`))
summary(test$autoid)
# BARBAR EPTNIL MYODAU NYCNOC   NoID  Noise PIPNAT PIPPYG PLEAUR VESMUR 
#     99  10924      1    177   1079  38532   1181   2676     19    230

# CM-49 
inputCM49 <- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-49/WAV/KPRO_V1/id.csv") 
# 22690 obs of 44 vars
# Add a site column 
inputCM49$site <- "CM-49"

## QUALITY CONTROL CHECK 

# Does every input file for each site correspond to only ONE detector? check serial number in the file names
## Manually checked - looks good! 

cm <- bind_rows(inputCM18, inputCM19, inputCM20,
                inputCM22, inputCM26, inputCM28, 
                inputCM39, inputCM40, inputCM49)  

29749 + 16684 + 19711 + 
44375 + 33595 + 54918 + 
14716 + 25388  + 22690 

dim(cm)
# 261826      of 45 vars
summary(cm) # No weird/obvious NAs 
#write.csv(cm, file = file.path(output_today, "AllSaraSites_MetaData_wCM26.csv")) 
# 13.01.2025

```

## Get an overview
```{r}

## Housekeeping
cm1 <- cm %>% 
  rename(
  filename = "OUT FILE FS",
  autoid = "AUTO ID*",
  DATE.12 = 'DATE-12',
  HOUR.12 = 'HOUR-12',
  TIME.12 = 'TIME-12') %>% 
  mutate(autoid = factor(autoid), 
         site = factor(site)) %>% 
  dplyr::select(OUTDIR, FOLDER, filename, DURATION, 
                DATE, TIME, HOUR,
                DATE.12, TIME.12, HOUR.12,
                autoid, site
                )
# Get a sense for how much noise there is across these sites 

## Make a figure illustrating this 
barnoise <- ggplot(cm1) + 
  geom_bar(aes(x= site, fill = autoid), position = "fill") +
  scale_fill_manual(name = "AutoID", values = moma.colors("Warhol", n = 13)) + 
  ylab("Proportion of recordings") + 
  xlab("Site") +
  theme(text = element_text(size = 25)) 
barnoise

# ggsave(path = output_today, filename = "AllSites_Sara_Noise_wCM26.tiff", width = 23, height = 14, device='tiff', dpi=300) 

#Very quick and dirty overview 

easytable <- table(cm1$site, cm1$autoid)

# write.csv(easytable, file = file.path(output_today, "EasySiteAutoIDOverview_wCM26.csv"))
# 
```


        BARBAR EPTNIL MYOBRA MYODAU MYOMYS MYONAT NYCNOC  NoID Noise
  CM-18     11   1356      0      2      0      0     14   147 27882
  CM-19     98   5682      0      0      0      0     13   545  9733
  CM-20     85  11010      0      0      0      0     50   414  7719
  CM-22    273  15143      0     15      1     15    512  2920 22832
  CM-26     56  22706      0      9      0      3    193  2068  7519
  CM-28     99  10924      0      1      0      0    177  1079 38532
  CM-39     47   1874      0      0      1      0      0   247 12531
  CM-40     33   7196     81     39    283      7     22  1026 16641
  CM-49    103   5785      0      3      0      4    159   961 14489
       
        PIPNAT PIPPYG PLEAUR VESMUR
  CM-18     93    221      5     18
  CM-19    463    140      3      7
  CM-20     54    211    158     10
  CM-22    353   1379      8    924
  CM-26    210    549     13    269
  CM-28   1181   2676     19    230
  CM-39     10      2      0      4
  CM-40     23      5      2     30
  CM-49    132    817     22    215


## Look at spatial temporal coverage 

```{r}
## Look at overall recording period 

ggplot(cm1) + 
  geom_point(aes(x = DATE.12, y = site), 
             color = "orange", alpha = 0.5, size = 3) +
    xlab("Date in season") + ylab(" ") +
  ggtitle("Recording activity") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        text = element_text(size = 25),
        axis.line = element_line(colour = "black"))

ggsave(path = output_today, filename = "BasicRecordingOverview_wCM26.tiff", width = 23, height = 14, device='tiff', dpi=300) 


## Good coverage! 

## Need to drop the weird early recording for CM-22 and the late recording for CM-39!

```



```{r}
## Remove noise and take another look 

levels(cm1$autoid)
summary(cm1$autoid)
# 157878 noise files that we will remove. 

cm2 <- cm1 %>% dplyr::filter(autoid != "Noise") %>% droplevels()
# 103948 obs of 12 vars 
# 261826 - 157878 = 103948- good!

## Simple dataset for easily making exploratory figures 
#write.csv(cm2, file = file.path(output_today, "SaraData_AutoId_NoNoise_SimpleColumns.csv"))

## Now take a look on what the recording activity looks like 
ggplot(cm2) + 
  geom_point(aes(x = DATE.12, y = site), 
             color = "turquoise", alpha = 0.5, size = 3) +
    xlab("Date in season") + ylab(" ") +
  ggtitle("Recording activity - Noise files removed") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        text = element_text(size = 25),
        axis.line = element_line(colour = "black"))

# ggsave(path = output_today, filename = "BasicRecordingOverview_noNoise.tiff", width = 23, height = 14, device='tiff', dpi=300) 


## And just because I am curious 
ggplot(cm2 %>% filter(autoid == "PIPNAT") %>% droplevels()) + 
  geom_point(aes(x = DATE.12, y = site), 
             color = "coral", alpha = 0.5, size = 3) +
    xlab("Date in season") + ylab(" ") +
  ggtitle("Recording activity - auotid Nathusius Pip") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        text = element_text(size = 25),
        axis.line = element_line(colour = "black"))

# ggsave(path = output_today, filename = "BasicRecordingOverview_PNATonly.tiff", width = 23, height = 14, device='tiff', dpi=300) 

```


## Finding a suitable subset that adds up to 20,000 files 
```{r}
## Need to drop the weird early recording for CM-22 and the late recording for CM-39!

levels(cm2$autoid)

DF <- cm2 %>% filter(
  autoid %in% c("BARBAR", "NYCNOC", "PIPNAT",
                "PIPPYG", "PLEAUR", "NoID")) %>% 
    droplevels() 

dim(DF)
# 20101    12  
# 13.01.2025
#write.csv(DF, file = file.path(output_today, "SaraSimpleColumn_Subset.csv"))
```

