---
title: "OverviewAutoProcessedData_PrepManualAnalysis__Benedikte2025"
output: html_document
date: "2025-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prepare and overview automatically classified bat acoustic data for Benedikte Øyens Masters thesis. 

## Combine AutoID .csv files from all Gismarvik and Utsira sites (CM and Noctur combined)
- Column for site and for detector type
## Check for gaps in data - get a sense for low activity periods and equipment failures 




## Prepare workspace
```{r}
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
library(janitor)
library(renv)
library(stringr)
library(beepr)
library(randomcoloR)
library(wesanderson)
library(leaflegend)
library(osmdata)
library(MetBrewer)
library(colorBlindness)
library(colorblindcheck)
library(MoMaColors)

getwd()
#  "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/GitHubLink/NocturBats"

## Setup output directory 
output <-"C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Analyses/Outputs/Reed" # where you want to save your data

file.name <- "OverviewProcessedBatData_Benedikte2025"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Analyses/Outputs/Reed/OverviewProcessedBatData_Benedikte2025_2025-01-09"

```

## Read in the processed data and look for potential errors 
```{r}

### NOCTUR DETECTORS - SM4-BATFS ### 

#Utsira
inputUT01 <- read_csv("P:/Noctur/DataCollection_2024/Acoustics/Utsira/UT-01/WAV/KPRO_NOC1/id.csv")
# 18209 obs of 44 vars
# Add a site column 
inputUT01$site <- "UT-01"

## Gismarvik
inputGI01 <- read_csv("P:/Noctur/DataCollection_2024/Acoustics/Gismarvik/GI-01/WAV/KPRO_NOC1/id.csv")
# 13623 obs of 44 vars
# Add a site column 
inputGI01$site <- "GI-01"

inputGI03 <- read_csv("P:/Noctur/DataCollection_2024/Acoustics/Gismarvik/GI-03/WAV/KPRO_NOC1/id.csv")
# 3025 obs of 44 vars 
# Add a site column 
inputGI03$site <- "GI-03"

## check column names

names(inputUT01) 
names(inputGI01)
names(inputGI03) 

## First glance - syntax looks consistent - cool!!
temp <- merge(inputGI01, inputGI03, all = TRUE) 
dim(temp)
# 16648  45
# 13623 + 3025 = 16648 - good! 

noc <- merge(temp, inputUT01, all = TRUE) 
dim(noc)
# 34857    45
# 18209 + 13623 + 3025 = 34857 - good!

## Now create a detector type column 

noc$detector <- "SM4-BATFS"
# 34857 46 

### Coastal Monitoring Detectors - Song Meter Mini Bat 2 Li-ion ### 

## Utsira 
inputCM01<- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-01/WAV/KPRO_V1/id.csv") 
# 24365 obs of 44 vars
# Add a site column 
inputCM01$site <- "CM-01"

inputCM02<- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-02/WAV/KPRO_V1/id.csv") 
# 7220 obs of 44 vars
# Add a site column 
inputCM02$site <- "CM-02"

## Gismarvik
inputCM46<- read_csv("P:/SW_CoastalMonitoring/Data_collection_2024/CM-46/WAV/KPRO_V1/id.csv") 
# 15304 obs of 44 vars
# Add a site column 
inputCM46$site <- "CM-46"

names(inputCM01)
names(inputCM02)
names(inputCM46)

# dim(inputCM01) + dim(inputCM02) = 31585
temp1 <- merge(inputCM01, inputCM02, all = TRUE) 

cm <- merge(temp1, inputCM46, all = TRUE)
# 31585 + 15304
# 46889 
# The number of rows add up - good! 

cm$detector <- "Song Meter Mini Bat 2 Li-ion"

## Now combine cm and noc

df <- merge(cm, noc, all = TRUE) 
dim(df)
# 81746    46

dim(cm) + dim(noc) # this matches, good!

# write.csv(df, file = file.path(output_today, "Gismarvik_Utsira_2024_raw_forBenedikte.csv"), row.names = FALSE)


```


## Exploring Benedikte's raw dataset 
```{r}

## Housekeeping
df1 <- df %>% 
  rename(
  filename = "OUT FILE FS",
  autoid = "AUTO ID*",
  DATE.12 = 'DATE-12',
  HOUR.12 = 'HOUR-12',
  TIME.12 = 'TIME-12') %>% 
  mutate(autoid = factor(autoid), 
         site = factor(site), 
         detector = factor(detector)) %>% 
  dplyr::select(OUTDIR, FOLDER, filename, DURATION, 
                DATE, TIME, HOUR,
                DATE.12, TIME.12, HOUR.12,
                autoid, site, detector
                )
  
summary(df1)

#What is the percentage of noise recorded across sites?
cm01 <- df1 %>% filter(site == "CM-01")  
cm02 <- df1 %>% filter(site == "CM-02")  
cm46 <- df1 %>% filter(site == "CM-46")  
ut01 <- df1 %>% filter(site == "UT-01")  
gi01 <- df1 %>% filter(site == "GI-01")  
gi03 <- df1 %>% filter(site == "GI-03")  
  
noiserat_cm01 <- cm01 %>% 
  group_by(autoid) %>%
  summarise(percentage = n() / nrow(cm01) * 100)
## 88% noise

noiserat_cm02 <- cm02 %>% 
  group_by(autoid) %>%
  summarise(percentage = n() / nrow(cm02) * 100)
## 98% noise

noiserat_cm46 <- cm46 %>% 
  group_by(autoid) %>%
  summarise(percentage = n() / nrow(cm46) * 100)
## 70% noise

noiserat_ut01 <- ut01 %>% 
  group_by(autoid) %>%
  summarise(percentage = n() / nrow(ut01) * 100)
## 96% noise

noiserat_gi01 <- gi01 %>% 
  group_by(autoid) %>%
  summarise(percentage = n() / nrow(gi01) * 100)
## 49% noise

noiserat_gi03 <- gi03 %>% 
  group_by(autoid) %>%
  summarise(percentage = n() / nrow(gi03) * 100)
## 55% noise
library(MoMAColors)

## Make a figure illustrating this 
barnoise <- ggplot(df1) + 
  geom_bar(aes(x= site, fill = autoid), position = "fill") +
  scale_fill_manual(name = "AutoID", values = moma.colors("Warhol", n = 13)) + 
  ylab("Proportion of recordings") + 
  xlab("Site") +
  theme(text = element_text(size = 25)) 
barnoise

ggsave(path = output_today, filename = "AllSites_Benedikte_Noise.tiff", width = 23, height = 14, device='tiff', dpi=300) 

## Look at overall recording period 

ggplot(df1) + 
  geom_point(aes(x = DATE.12, y = site), 
             color = "orange", alpha = 0.5, size = 3) +
    xlab("Date in season") + ylab(" ") +
  ggtitle("Recording activity") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        text = element_text(size = 25),
        axis.line = element_line(colour = "black"))

ggsave(path = output_today, filename = "BasicRecordingOverview.tiff", width = 23, height = 14, device='tiff', dpi=300) 

# My first instinct is that most recording across all sites happened between late July and early September


```


## Remove noise and take another look 

```{r}
levels(df1$autoid)
summary(df1$autoid)
# 65226     noise files that we will remove. 

df2 <- df1 %>% dplyr::filter(autoid != "Noise") %>% droplevels()
# 16520 obs of 13 vars 
# 81746 - 65226 = 16520 - good! 

## Simple dataset for easily making exploratory figures 
#write.csv(df2, file = file.path(output_today, "BenedikteData_AutoId_NoNoise_SimpleColumns.csv"))

## Now take a look on what the recording activity looks like 
ggplot(df2) + 
  geom_point(aes(x = DATE.12, y = site), 
             color = "turquoise", alpha = 0.5, size = 3) +
    xlab("Date in season") + ylab(" ") +
  ggtitle("Recording activity - Noise files removed") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        text = element_text(size = 25),
        axis.line = element_line(colour = "black"))

ggsave(path = output_today, filename = "BasicRecordingOverview_noNoise.tiff", width = 23, height = 14, device='tiff', dpi=300) 


## And just because I am curious 
ggplot(df2 %>% filter(autoid == "PIPNAT") %>% droplevels()) + 
  geom_point(aes(x = DATE.12, y = site), 
             color = "coral", alpha = 0.5, size = 3) +
    xlab("Date in season") + ylab(" ") +
  ggtitle("Recording activity - auotid Nathusius Pip") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        text = element_text(size = 25),
        axis.line = element_line(colour = "black"))

ggsave(path = output_today, filename = "BasicRecordingOverview_PNATonly.tiff", width = 23, height = 14, device='tiff', dpi=300) 

### Now add a subset of random noise files from reach site for Benedikte to also check 

table(df1$site, df1$autoid)
# Randomly select 500 Noise files from each site for her to validate

noise_sub <- df1 %>% 
  filter(autoid == "Noise") %>% 
  droplevels() %>% 
  group_by(site) %>% slice_sample(n=500) %>% 
  ungroup() 
dim(noise_sub)
# 3000 13 

summary(noise_sub$site)
# CM-01 CM-02 CM-46 GI-01 GI-03 UT-01 
#   500   500   500   500   500   500 

## Tada!

## Now add these back to Benedikte main dataset 

df3 <- merge(df2, noise_sub, all = TRUE)
# The numbers of rows add up! 

# write.csv(df3, file = file.path(output_today, "CompleteDatasetBenedikte_wNoiseSubset_SimpleColumnsOnly.csv"))

```


