---
title: "SiteSelection_WindFarms2025"
output: html_document
date: "2025-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import the points of all the wind turbines for 
Høg-Jæren
Måkaknuten
Stigafjellet
Egersund 
Bjerkreim

### Explore variation in elevation across all points


```{r}
library(tidyverse)
library(dplyr)
library(raster)
library(rgdal)
library(rgeos)
library(mapview)
library(sf)
library(sp)
library(maps)
library(leaflet)
library(randomcoloR)
library(wesanderson)
library(leaflegend)
library(osmdata)
library(MetBrewer)
library(colorBlindness)
library(colorblindcheck)
library(ggmap)
library(maptools)
library(spatstat)
library(elevatr)
library(gdalUtils)
library(rgdal)
library(terra)


getwd()
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/GitHubLink/NocturBats"

## Setup output directory 

output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/GitHubLink/CoastalMonitoring/Outputs"

file.name <- "SiteSelection_WindFarms2025"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today

#"C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/GitHubLink/CoastalMonitoring/Outputs/SiteSelection_WindFarms2025_2025-02-17"

```


```{r}

## Note that I had to manually generate the points for Høg-Jæren so these are not named according to the actual turbine identities. 


## The KML file has the points separated into folder for each wind farm so I need to use a function to simulatenously import all layers at once 
allKmlLayers <- function(kmlfile){
  lyr <- ogrListLayers(kmlfile)
  mykml <- list()
  for (i in 1:length(lyr)) {
    mykml[i] <- readOGR(kmlfile,lyr[i])
  }
  names(mykml) <- lyr
  return(mykml)
}


kmlfile <- ("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Locations/StudyDesign_NorskVind_WTs_2025.KML")

tpt <- allKmlLayers(kmlfile)

## Just for the Høg-Jæren study design proposal
hjsd <- st_read("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Locations/HJ_SiteSelectionPt12025.KML")

## This is how you import the folders one at a time: 

bj <- readOGR(kmlfile,"Bjerkreim") # 70 pts
eg <- readOGR(kmlfile,"Egersund") # 33 pts
ma <- readOGR(kmlfile,"Maakaknuten") # 22 pts
st <- readOGR(kmlfile,"Stigafjellet") # 7 pts 
hj <- readOGR(kmlfile,"Hoeg-Jaeren") # 32 pts
## Number of points correspond to what is expected 
## Add a WindFarm column 

bj$WindFarm <- "Bjerkreim"
eg$WindFarm <- "Egersund"
ma$WindFarm <- "Maakaknuten"
st$WindFarm <- "Stigafjellet"
hj$WindFarm <- "Hoeg-Jaeren"

## Right now the names for Egersund and Bjerkreim wind turbines are not terribly helpful so I am going to fix those

bj$Name <- gsub("E", "B", bj$Name)
bj$Name <- gsub("S", "B", bj$Name)
bj$Name <- gsub("G", "B", bj$Name)

eg$Name <- gsub("S", "E", eg$Name)

crs(bj)
crs(eg)
crs(ma)
crs(st)
crs(hj)

crs(hjsd)
## All 4326


## Now try to combine everyone 
temp1 <- spRbind (bj, eg) # 103 pts
temp2 <- spRbind(temp1, ma) # 125 pts 
temp3 <- spRbind(temp2, st) # 132 pts 
temp4 <- spRbind(temp3, hj) # 164 pts 
# 70 + 33 + 22 + 7 + 32
# 164 - good ! 

crs(temp4)

tpt <- temp4 
tpt$WindFarm <- factor(tpt$WindFarm)
levels(tpt$WindFarm)

## Now plot it to see if it makes sense 
plot(tpt)

# Prettier plot
test <- MetBrewer::scale_color_met_d("Java")

pal <- colorFactor(
  c("#663171","#e2998a","#cf3a36","#0c7156","#ea7428"), 
  domain = c("Bjerkreim",    "Egersund",     "Hoeg-Jaeren",  "Maakaknuten",  "Stigafjellet")) 

plot1 <- tpt %>% leaflet() %>% 
   addTiles()  %>% 
   addCircleMarkers(radius = 3, color = ~pal(WindFarm), 
                       fillOpacity = 1,  stroke = TRUE) %>% 
  addMiniMap(
    position = "bottomright", zoomLevelOffset = -7, width = 300, height = 300, 
    collapsedWidth = 19, collapsedHeight = 19) %>% 
  addScaleBar(position = "bottomright")
plot1$x$options = append(plot1$x$options, list("zoomControl" = FALSE,  "scaleBar"=TRUE))
plot1




## Add site type column 
hjsd$Name

hjsd1 <- hjsd %>% 
  mutate(SiteType = factor(case_when(
  Name %in% c("HO-03",       "HO-09",
              "HO-24",       "HO-31",
              "HO-27",       "HO-22") ~ "Primary", 
  TRUE ~ "Control"
)))

## Just so it is easier to plot
hjsd1$Name
hjsd1$Name <- gsub("CT-24b", "CT-24b   ", hjsd1$Name)
hjsd1$Name <- gsub("CT-31a", "CT-31a   ", hjsd1$Name)
hjsd1$Name

summary(hjsd1$SiteType)
# Control Primary 
#       8       6

pal <- colorFactor(
  c("black", "#CF3A36"), 
  domain = c("Primary",    "Alternative Primary",     "Control")) 

plot1 <- hjsd1 %>% leaflet() %>% 
   addTiles()  %>% 
   addCircleMarkers(radius = 9, color = ~pal(SiteType), 
                       fillOpacity = 1,  stroke = FALSE)
plot1

##  addMiniMap(
#     position = "bottomright", zoomLevelOffset = -4, width = 300, height = 300, 
#     collapsedWidth = 19, collapsedHeight = 19) %>% 
#   addScaleBar(position = "bottomright")
# plot1$x$options = append(plot1$x$options, list("zoomControl" = FALSE,  "scaleBar"=TRUE))

# 
# ,
#                     label = ~Name,
#   labelOptions = labelOptions(noHide = TRUE, offset=c(2,-9), textOnly = TRUE, textsize = "15px")) 
# ```


## Extract elevation and explore
https://cran.r-project.org/web/packages/elevatr/vignettes/introduction_to_elevatr.html
- This only really works for coordinates inside the U.S. unfortunately

Got some help from Jaime instead
https://rdrr.io/cran/terra/man/extract.html
r <- rast(path_to_tif)
p <- vect(path_to_points)
p <- project(p, crs(r, proj=TRUE))

p_heiths = extract(p, r)



```{r}

tpt1 <- sf::st_as_sf(tpt, coords = c("x", "y"), crs = 4326)

writeOGR(tpt1, dsn = '.', layer = 'poly', driver = "ESRI Shapefile")

# Combine raster data
## Following example from here 
## https://stackoverflow.com/questions/15876591/merging-multiple-rasters-in-r

r1 <- raster("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Locations/DTM_Elevation_StudyArea_Geonorge/6403_1_10m_z32.tif")

r2 <- raster("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Locations/DTM_Elevation_StudyArea_Geonorge/6503_2_10m_z32.tif")

r3 <- raster("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Locations/DTM_Elevation_StudyArea_Geonorge/6503_3_10m_z32.tif")

r4 <- raster("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Locations/DTM_Elevation_StudyArea_Geonorge/6403_4_10m_z32.tif")

tmp <- raster::merge(r1, r2)
tmp1 <- raster::merge(tmp, r3)
tmp2 <- raster::merge(tmp1, r4)

r <-  as(tmp2, "SpatRaster")
p <- vect(tpt1)
p <- project(p, crs(r, proj=TRUE)) 

## Elevation along with the coordinates of each point - now to merge with the points 
p_height = terra::extract(r, p, ID = FALSE, xy = TRUE, bind = TRUE)

## Now we can just look at the elevation values but cannot plot the data in a map
DF <- as.data.frame(p_height) %>% 
  rename(elevation = layer) %>% arrange(WindFarm, elevation)

## Now make a simpler wind turbine ID column 

### Now look for differences in elevation across the different wind turbines in each wind farm

ggplot(DF %>% arrange(WindFarm, elevation)) +
  geom_col(aes(x=Name, y = elevation, fill = WindFarm), position = "dodge") + 
  scale_fill_manual(values = c("#663171","#e2998a","#cf3a36","#0c7156","#ea7428"), name = "") + 
  theme_minimal() + 
  xlab("Wind turbines") + 
  ylab("Elevation (meters of sea level)")

source("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/BatLab Norway/Projects/Noctur/Analyses/group_by_summary_stats.R")
  
group_by_summary_stats(DF, elevation, WindFarm)

#   WindFarm         N  Mean Median    SD    SE Range      
#   <fct>        <int> <dbl>  <dbl> <dbl> <dbl> <chr>      
# 1 Bjerkreim       70  258.   250.  38.5  4.60 183.4-348.4
# 2 Egersund        33  248.   247.  36.1  6.29 175.5-330.8
# 3 Hoeg-Jaeren     32  233.   234.  22.5  3.98 192.2-268.3
# 4 Maakaknuten     22  374.   370.  28.7  6.11 320.3-420.8
# 5 Stigafjellet     7  399.   400.  16.9  6.40 375.1-424.4

table(DF$Name, DF$elevation)


bj <- DF %>% filter(WindFarm == "Bjerkreim") %>% droplevels()
summary(bj$elevation)

ggplot(bj %>% arrange(WindFarm, elevation)) +
  geom_col(aes(x=reorder(Name, elevation), y = elevation, fill = WindFarm), position = "dodge") + 
  scale_fill_manual(values = c("#663171"), name = "") + 
  theme_minimal() + 
  xlab("Wind turbines") + 
  ylab("Elevation (meters of sea level)")


hj <- DF %>% filter(WindFarm == "Hoeg-Jaeren") %>% droplevels()
summary(hj$elevation)

ggplot(hj %>% arrange(WindFarm, elevation)) +
  geom_col(aes(x=reorder(Name, elevation), y = elevation, fill = WindFarm), position = "dodge") + 
  scale_fill_manual(values = c("#cf3a36"), name = "") + 
  theme_minimal() + 
  xlab("Wind turbines") + 
  ylab("Elevation (meters of sea level)")


eg <- DF %>% filter(WindFarm == "Egersund") %>% droplevels()
summary(eg$elevation)

ggplot(eg %>% arrange(WindFarm, elevation)) +
  geom_col(aes(x=reorder(Name, elevation), y = elevation, fill = WindFarm), position = "dodge") + 
  scale_fill_manual(values = c("#e2998a"), name = "") + 
  theme_minimal() + 
  xlab("Wind turbines") + 
  ylab("Elevation (meters of sea level)")


ma <- DF %>% filter(WindFarm == "Maakaknuten") %>% droplevels()
summary(ma$elevation)


ggplot(ma %>% arrange(WindFarm, elevation)) +
  geom_col(aes(x=reorder(Name, elevation), y = elevation, fill = WindFarm), position = "dodge") + 
  scale_fill_manual(values = c("#0c7156"), name = "") + 
  theme_minimal() + 
  xlab("Wind turbines") + 
  ylab("Elevation (meters of sea level)")


st <- DF %>% filter(WindFarm == "Stigafjellet") %>% droplevels()
summary(st$elevation)


ggplot(st %>% arrange(WindFarm, elevation)) +
  geom_col(aes(x=reorder(Name, elevation), y = elevation, fill = WindFarm), position = "dodge") + 
  scale_fill_manual(values = c("#ea7428"), name = "") + 
  theme_minimal() + 
  xlab("Wind turbines") + 
  ylab("Elevation (meters of sea level)")



```


## Create radii around selected wind turbines to estimate areas for deploying equipment 
